---
title: "NASDAQ"
output: html_document
---

I chose Cisco Systems, Inc. (CSCO) from NASDAQ. This anyalaiss focuses on  close price and volume The data was downloaded from http://finance.yahoo.com/q/hp?s=CSCO+Historical+Prices on January 31, 2016. 
```{r echo=FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(pander)
library(stats)
library(xts)
library(sandwich)

# Open
src = "http://real-chart.finance.yahoo.com/table.csv?s=CSCO&d=0&e=31&f=2016&g=d&a=2&b=26&c=1990&ignore=.csv"
src = "csco.csv" # Comment this line in order to download the latest data from web
cisco <- read.csv(src, stringsAsFactors = FALSE)
# Clean and Transform
setDT(cisco)
setkey(cisco, Date)
cisco[, Date := as.Date(Date)]
cisco[, Open := NULL]
cisco[, High := NULL]
cisco[, Low := NULL]
cisco[, Adj.Close := NULL]
```

# Data
```{r echo=FALSE, warning=FALSE, message=FALSE}
# Explore
pander(summary(cisco))
ggplot(cisco, aes(x = "Close", y = Close)) + geom_boxplot() + ggtitle("Variance of Close price") + theme_minimal()
ggplot(cisco, aes(x = Close)) + geom_histogram() + ggtitle("Histogram of Close price") + theme_minimal()
ggplot(cisco, aes(x = "Volume", y = Volume)) + geom_boxplot() + ggtitle("Variance of Volume") + theme_minimal()
ggplot(cisco, aes(x = Volume)) + geom_histogram() + ggtitle("Histogram of Volume") + theme_minimal()
ggplot(cisco, aes(x = Date, y = Close)) + geom_line() + ggtitle("Close price over time") + theme_minimal()
ggplot(cisco, aes(x = Date, y = Volume)) + geom_line() + ggtitle("Volume over time") + theme_minimal()
```

# Close Price
## Stationarity
The "Close price over time" charts seems to show the characteristirics of a random walk. But the Phillips-Perron test refuses with a high confidence the hypothesis that it has a unit root.
```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Close))
```

Although the original series can be accepted as stationary with high confidence, further improvements can be achieved via transforming the close price: taking the log difference results in a more stationary model. 
```{r echo=FALSE, warning=FALSE}
cisco[, Close.Log := log(Close)]
cisco[, Close.Log.Diff := Close.Log - shift(Close.Log, n=1, fill=NA, type="lag")]
ggplot(cisco, aes(x = Date, y = Close.Log.Diff)) + geom_line() + 
  ggtitle("Log difference of Close price over time") + theme_minimal()
```
```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Close.Log.Diff[-1]))
```

## OLS model with Newey-West SE using 2 lags
```{r echo=FALSE, warning=FALSE}
cisco[, Close.Log.Diff.lag.1 := shift(Close.Log.Diff, n=1, fill=NA, type="lag")]
cisco[, Close.Log.Diff.lag.2 := shift(Close.Log.Diff, n=2, fill=NA, type="lag")]
fit <- lm(Close.Log.Diff ~ Close.Log.Diff.lag.1 + Close.Log.Diff.lag.2, data = cisco)
se <- sqrt(diag(NeweyWest(fit)))
df <- data.frame()
df <- rbind(df,fit$coefficients,se)
colnames(df) <- labels(fit$coefficients)
rownames(df) <- c("Coefficients", "Newey-West SE")
pander(df)
```

Based on OLS with Newey-West SE using 2 lags the returns can not be predicted. It is an evidence of Efficient Market Hypotheses

## Monday Effect
Do you see evidence of Monday effect in returns?

## Time Periods 
Compare two time periods of your choice. Do you see any important  difference?

# Volume
## Stationarity
The "Volume over time" charts seems to be stationary. Also, the Phillips-Perron test refuses with a high confidence the hypothesis that it has a unit root.
```{r echo=FALSE, warning=FALSE}
pander(PP.test(as.numeric(cisco$Volume)))
```

Although the original series can be accepted as stationary with high confidence, further improvements can be achieved via transforming the volume: taking the log difference results in a more stationary model. 
```{r echo=FALSE, warning=FALSE}
cisco[, Volume.Log := log(Volume)]
cisco[, Volume.Log.Diff := Volume.Log - shift(Volume.Log, n=1, fill=NA, type="lag")]
ggplot(cisco, aes(x = Date, y = Volume.Log.Diff)) + geom_line() + 
  ggtitle("Log difference of Volume over time") + theme_minimal()
```
```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Volume.Log.Diff[-1]))
```

## Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
volumes <- xts(cisco$Volume.Log.Diff[-1], cisco$Date[-1])
acf(volumes)
```

## Partial Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
pacf(volumes)
```

## ARMA(2,2)  
```{r echo=FALSE, warning=FALSE}
model <- arima(volumes, c(2,0,2))
pander(model)
```

## Change in Trend
Do you see a change in trend for the volume?


# Regression

Consider these two models

(1) E lnSIt

(2) E lnSIt

where RS is the return of your stock, RI is the return of nasdaq composite, si

are seasonal dummies.

1. Run an OLS regressions with newey-west SE with two lags for each 

2. Present results in one table.

3. Consider the first model. Interpret the β coefficient in words. What 

4. For extra credit: repeat this with ARMA(2,2) and discuss stability of 

models. 

economic/finance concept could you relate it to?

results.
