---
title: "NASDAQ"
output: html_document
---

I chose Cisco Systems, Inc. (CSCO) from NASDAQ. This analysis focuses on close price and volume. The data was downloaded from http://finance.yahoo.com/q/hp?s=CSCO+Historical+Prices on January 31, 2016.

```{r echo=FALSE, warning=FALSE, message=FALSE}
library(data.table)
library(ggplot2)
library(pander)
library(stats)
library(xts)
library(sandwich)

# Open
src = "http://real-chart.finance.yahoo.com/table.csv?s=CSCO&d=0&e=31&f=2016&g=d&a=2&b=26&c=1990&ignore=.csv"
src = "csco.csv" # Comment this line in order to download the latest data from web
cisco <- read.csv(src, stringsAsFactors = FALSE)
# Clean and Transform
setDT(cisco)
setkey(cisco, Date)
cisco[, Date := as.Date(Date)]
cisco[, Open := NULL]
cisco[, High := NULL]
cisco[, Low := NULL]
cisco[, Adj.Close := NULL]
```

# Data

```{r echo=FALSE, warning=FALSE, message=FALSE, fig.width=3.5, fig.height=2.6}
# Explore
pander(summary(cisco))
ggplot(cisco, aes(x = "Close", y = Close)) + geom_boxplot() + ggtitle("Variance of Close price") + theme_minimal()
ggplot(cisco, aes(x = Close)) + geom_histogram() + ggtitle("Histogram of Close price") + theme_minimal()
ggplot(cisco, aes(x = "Volume", y = Volume)) + geom_boxplot() + ggtitle("Variance of Volume") + theme_minimal()
ggplot(cisco, aes(x = Volume)) + geom_histogram() + ggtitle("Histogram of Volume") + theme_minimal()
ggplot(cisco, aes(x = Date, y = Close)) + geom_line() + ggtitle("Close price over time") + theme_minimal()
ggplot(cisco, aes(x = Date, y = Volume)) + geom_line() + ggtitle("Volume over time") + theme_minimal()
```

# Close Price
## Stationarity
The "Close price over time" charts seems to show the characteristics of a random walk. But the Phillips-Perron test rejects the hypothesis that it has a unit root with a high confidence.

```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Close))
```

Although the original series can be accepted as stationary with high confidence, further improvements can be achieved via transforming the close price: taking the log difference results in a "more" stationary model. 

```{r echo=FALSE, warning=FALSE}
cisco[, Close.Log := log(Close)]
cisco[, Close.Log.Diff := Close.Log - shift(Close.Log, n=1, fill=NA, type="lag")]
ggplot(cisco, aes(x = Date, y = Close.Log.Diff)) + geom_line() + 
  ggtitle("Log difference of Close price over time") + theme_minimal()
```
```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Close.Log.Diff[-1]))
```

## OLS model with Newey-West SE using 2 lags
Based on OLS with Newey-West SE using 2 lags the returns can not be predicted. It is an evidence of Efficient Market Hypotheses.

```{r echo=FALSE, warning=FALSE}
cisco[, Close.Log.Diff.lag.1 := shift(Close.Log.Diff, n=1, fill=NA, type="lag")]
cisco[, Close.Log.Diff.lag.2 := shift(Close.Log.Diff, n=2, fill=NA, type="lag")]
fit <- lm(Close.Log.Diff ~ Close.Log.Diff.lag.1 + Close.Log.Diff.lag.2, data = cisco)
se <- sqrt(diag(NeweyWest(fit, lag = 2)))
df <- data.frame()
df <- rbind(df,fit$coefficients,se)
colnames(df) <- labels(fit$coefficients)
rownames(df) <- c("Coefficients", "Newey-West SE")
pander(df)
```

## Monday Effect
Based on OLS with Newey-West SE using Monday dummy variable there is no evidence for Monday effect when considering the close prices between 1990 and 2016.

```{r echo=FALSE, warning=FALSE}
cisco[, Weekday := weekdays(Date)]
cisco[, Weekday.Monday := Weekday == "Monday"]
fit <- lm(Close.Log.Diff ~ Weekday.Monday, data = cisco)
se <- sqrt(diag(NeweyWest(fit, lag = 2)))
df <- data.frame()
df <- rbind(df,fit$coefficients,se)
colnames(df) <- labels(fit$coefficients)
rownames(df) <- c("Coefficients", "Newey-West SE")
pander(df)
```

## Time Periods
Based on the "Close price over time" chart there is a significant difference in close price before and after March, 2000. According to [Market Crashes: The Dotcom Crash](the http://www.investopedia.com/features/crashes/crashes8.asp) article, The Nasdaq Composite lost 78% of its value as it fell from 5046.86 to 1114.11. 

```{r echo=FALSE, warning=FALSE}
dotcom <- as.Date("2000-03-11")
cisco[, Dotcom.Before := Date <= dotcom ][,Dotcom.After := Date > dotcom]
ggplot(cisco, aes(x = Dotcom.After, y = Close)) + geom_boxplot()  + ggtitle("The Dotcom Crash of Close price") + theme_minimal()
```

# Volume
## Stationarity
The "Volume over time" charts seems to be stationary. Also, the Phillips-Perron test rejects the hypothesis that it has a unit root with a high confidence.

```{r echo=FALSE, warning=FALSE}
pander(PP.test(as.numeric(cisco$Volume)))
```

Although the original series can be accepted as stationary with high confidence, further improvements can be achieved via transforming the volume: taking the log difference results in a "more" stationary model. 

```{r echo=FALSE, warning=FALSE}
cisco[, Volume.Log := log(Volume)]
cisco[, Volume.Log.Diff := Volume.Log - shift(Volume.Log, n=1, fill=NA, type="lag")]
ggplot(cisco, aes(x = Date, y = Volume.Log.Diff)) + geom_line() + 
  ggtitle("Log difference of Volume over time") + theme_minimal()
```
```{r echo=FALSE, warning=FALSE}
pander(PP.test(cisco$Volume.Log.Diff[-1]))
```

## Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
volumes <- xts(cisco$Volume.Log.Diff[-1], cisco$Date[-1])
acf(volumes)
```

## Partial Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
pacf(volumes)
```

## ARMA(2,2)  
```{r echo=FALSE, warning=FALSE}
model <- arima(volumes, c(2,0,2))
pander(model)
```

## Change in Trend
Similarly to the close price, the Dotcom Crash had an effect on daily volumes as well. But this effect is not as significant as on the close price.

```{r echo=FALSE, warning=FALSE}
ggplot(cisco, aes(x = Dotcom.After, y = Volume)) + geom_boxplot()  + 
  ggtitle("The Dotcom Crash of Volume") + theme_minimal()
```

# Regression

(1) $E[ lnRS_{t} | lnRI_{t} ] = \alpha + \beta * RI_{t}$

(2) $E[ lnRS_{t} | lnRI_{t} ] = \alpha + \beta * RI_{t} + \gamma * t + sum ( O_{i} * s_{i} )$ 

where RS is the return of your stock, RI is the return of nasdaq composite, s are seasonal dummies.

```{r echo=FALSE, warning=FALSE}
# Download and calculate return on nasdaq composite
src = "http://real-chart.finance.yahoo.com/table.csv?s=%5EIXIC&a=02&b=26&c=1990&d=00&e=29&f=2016&g=d&ignore=.csv"
src = "nasdaq.csv"
nasdaq <- read.csv(src, stringsAsFactors = FALSE)
# Clean and Transform
setDT(nasdaq)
setkey(nasdaq, Date)
cisco[, Nasdaq.Close := nasdaq$Close]
cisco[, Nasdaq.Close.Log := log(Nasdaq.Close)]
cisco[, Nasdaq.Close.Log.Diff := Nasdaq.Close.Log - shift(Nasdaq.Close.Log, n=1, fill=NA, type="lag")]
```

## OLS model with Newey-West SE using 2 lags
```{r echo=FALSE, warning=FALSE}
# Model 1
model <- lm(Close.Log.Diff ~ Nasdaq.Close.Log.Diff, data = cisco)
se <- sqrt(diag(NeweyWest(model, lag = 2)))
dt1 <- data.table(labels(model$coefficients))
dt1 <- cbind(1:2, dt1, model$coefficients, se)
colnames(dt1) <- c("i", "var", "Coefficients (1)", "Newey-West SE (1)")

model <- lm(Close.Log.Diff ~ Nasdaq.Close.Log.Diff +
               Date + Weekday.Monday + Dotcom.Before, data = cisco)
se <- sqrt(diag(NeweyWest(model, lag = 2)))
dt2 <- data.table(labels(model$coefficients))
dt2 <- cbind(1:5, dt2, model$coefficients, se)
colnames(dt2) <- c("i", "var", "Coefficients (2)", "Newey-West SE (2)")

dt <- merge(dt1, dt2, by = "var", all.y = TRUE)
setorder(dt, i.y)
dt[, i.x := NULL][, i.y := NULL]
pander(dt, split.table = Inf)
```

## β coefficient 
The contemporaneous log difference of NASDAQ composite index has a significant coefficient of 1.318 considering the standard error of 0.026. It means that one unit difference in the NASDAQ composite index is expected to result in 1.3 unit difference in the Cisco stock price. So whenever the NASDAQ composite index is 10 percentage points higher than the previous day, the Cisco is expected to close with 13 percentage points higher price. 

The relative higher proportional increase of price could be explained by the relative strength of the Cisco stock. Cisco can be considered as a Blue-chip company. 

# Extra Credit: ARMA(2,2) 
The ARMA model is built on the time series of the Cisco stock price. The insignificance of the coefficients is the evidence of Efficient Market Hypotheses. 

## Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
prices <- xts(cisco$Close.Log.Diff[-1], cisco$Date[-1])
acf(prices)
```

## Partial Autocorrelation Function
```{r echo=FALSE, warning=FALSE}
pacf(prices)
```

## Model (1)
```{r echo=FALSE, warning=FALSE}

model <- arima(prices, c(2,0,2), xreg = cisco[-1,.(Nasdaq.Close.Log.Diff)])
pander(model)
```

## Model (2)
```{r echo=FALSE, warning=FALSE}
model <- arima(prices, c(2,0,2), xreg = cisco[-1,.(Nasdaq.Close.Log.Diff, Weekday.Monday, Dotcom.Before)])
pander(model)
```